---
title: "Practive descriptive"
output:
  html_document:
  toc: true
number_sections: true
author: "by: Stijn van Helmondt"
bibliography: references.bib

---

# Starting up
 
## Packages

```{r}
rm(list = ls())


```

## Functions

Explain what function does what
```{r}
# Loading in packages

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

## Packages

Tidyverse --> basic data manipulation

```{r}
packages = c("stringdist", "stringi", "tidyverse", "igraph")

fpackage.check(packages)
```


## Data input

Where data comes from. Comes from ... (website), could also be from own scraping. 
Short description of what is in the dataset

```{r}
df <- fload("./Data/Data_processed/soc_df.RData")
publications <- fload("./Data/Data_processed/soc_df_publications.RData")
```

Getting to know the data
```{r}
fshowdf(df)


summary(df$soc_names)

#mytable <- table(df$uni, useNA = c("always"))
#mytable

#mytable2 <- table(df$affil1, df$discipline, df$affil2)
#mytable2

#?table 

```

Cleaning the data 
```{r}
df$uni == "UU/VU"

#selecting a specific value
df$uni[df$uni == "UU/VU"]

#Recoding
df$uni2 <- df$uni

df$uni2[df$uni == "UU/VU"] <- "UU"



```

Only keep unique publications
```{r}

publications <- publications %>%
    bind_rows() %>%
    distinct(title, .keep_all = TRUE)
```

## Data output

```{r}
outputRU_sociology <- f_pubnets(affiliation = "RU")
str(outputRU_sociology)

outputRU_sociology$network[[1]]

wave1 <- outputRU_sociology$network[[1]]
wave2 <- outputRU_sociology$network[[2]]
```

Plotting within the raw data
```{r}
# let us put the diagonal to zero
diag(wave1) <- 0
diag(wave2) <- 0

wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
```

## Plotting wave 1
```{r}
G1 <- igraph::graph_from_adjacency_matrix(wave1, mode = "directed", weighted = TRUE, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

plot(G1)

edges <- igraph::as_data_frame(G1, what = "edges")
G1 <- graph_from_data_frame(edges, directed = TRUE)



```
Ei-index --> hoeveel within group ties 
Kijken naar verschil between group en ingroup, gedeeld door totaal aantal ties 
Alleen kijken naar bestaande ties!!! 
Wel vaak gebruikt voor egonets, niet voor een volledig netwerk, daarvoor werken proporties beter. 

Maat van segregatie die werkt met weighted ties. 

Een van die maten is Moran's I

Maat wordt positief als nodes op elkaar lijken 

```{r}

```


