---
title: "Descriptive statistics"
output:
  html_document:
    toc: true
    number_sections: true
author: "by: Stijn van Helmondt"
bibliography: references.bib
csl: apa.csl

---

# Starting up

## Loading in packages

```{r}
rm(list = ls())


```

## Functions

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

f_pubnets <- function(df_scholars = df, list_publications = publications,
    waves = list(wave1 = c(2018, 2019), wave2 = c(2020, 2021), wave3 = c(2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```


## Data input

Where data comes from. Comes from ... (website), could also be from own scraping. 
Short description of what is in the dataset

```{r}
df <- fload("./Data/Data_processed/soc_df.RData")
publications <- fload("./Data/Data_processed/soc_df_publications.RData")
```





# Taking a look at the data

```{r}
rm(list = ls())

#### packages ####.
require(tidyverse)
require(igraph)

##### Data input ###.
load("addfiles/soc_df.Rdata")
load("addfiles/soc_df_publications.Rdata")
load("addfiles/soc_collabs1.RData")
load("addfiles/soc_collabs2.RData")


summary(soc_df$soc_names)

str(soc_df)
view(soc_df_publications)

```

# Recoding variables
```{r}
soc_df$genderman <- NA
# fill category by category
soc_df$genderman[soc_df$gender == "man" ] <- 1
soc_df$genderman[soc_df$gender == "woman" ] <- 0

summary(soc_df$genderman)
```

# Making graphs
```{r}
# get unique collaborators of soc staff first
soc_collabs_unique <- unique(soc_collabs[, 4])
soc_collabs_unique <- c(soc_collabs_unique, soc_df$gs_id)  
soc_collabs_unique <- data.frame(soc_collabs_unique)
soc_collabs_unique$v1 <- 1  
soc_collabs_unique <- unique(soc_collabs_unique)

onefivedegree <- left_join(collabs_1deep, soc_collabs_unique, by = c(coauth_id = "soc_collabs_unique"))

# Drop those that are not among sociology collaborators and who don't list coauthors regardless
onefivedegree <- onefivedegree[!is.na(onefivedegree$v1), ]
onefivedegree <- onefivedegree[!is.na(onefivedegree$coauth), ]

# Pick those columns and have an edgelist of soc collaborators and whether they collaborate with those same collaborators the 1.5 degree network
onefivedegree <- onefivedegree[, c("name", "coauth")]
names(onefivedegree) <- c("from", "to")

# we get soc collaborators and add those to the data above and end up with a nice edgelist!
socc <- soc_collabs[!is.na(soc_collabs$coauth), ]
socc <- socc[, c("name", "coauth")]
names(socc) <- c("from", "to")
onefivedegree <- rbind(socc, onefivedegree)
save(onefivedegree, file = "addfiles/soc_onefivedegree.RData")
```


# Plotting
```{r}
onefivedegree <- as.matrix(onefivedegree) 
net1 <- graph_from_edgelist(onefivedegree, directed = TRUE) 
plot(net1)
```
# Making the graph prettier

```{r}
plot(net1 ,
     vertex.color = "gold", # nice color for the vertices
     vertex.size = 4,  # we'll vertices a bit smaller
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.family = "Helvetica", # not a fan of times new roman in figures
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1) # make arrow size (direction of edge) smaller


in_network <- data.frame(as_ids(V(net1)))
names(in_network)[1] <- "name"

ru_nodes <- soc_df[, c("name", "i10_index")]
in_network <- left_join(in_network, ru_nodes, by = c("name" = "name"))
in_network$vcol <- ifelse(is.na(in_network$i10_index), "#E69F00", "#56B4E9")

plot(net1,
     vertex.color = in_network$vcol, #THIS WAS WHAT WE DID THE LAST CODEBLOCK FOR!
     vertex.size = 4,  # we'll make them a bit smaller
     vertex.frame.color = "gray",  # we'll put a frame around it
     vertex.label.color = "black",  # not that ugly blue color for the names
     vertex.label.family = "Helvetica", # not a fan of times in figures
     vertex.label.cex = 0.4,  # a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2,  # curved edges is always a nice tough
     edge.arrow.size = 0.1) # arrow size smaller
```

# Only names RU staff
```{r}
plot(net1,
     vertex.color = in_network$vcol, 
     vertex.label = ifelse(!is.na(in_network$i10_index), in_network$name, NA),
     vertex.size = 5,  # we'll make them a bit smaller
     vertex.frame.color = "gray",  # we'll put a frame around it
     vertex.label.color = "black",  # not that ugly blue color for the names
     vertex.label.family = "Helvetica", # not a fan of times in figures
     vertex.label.cex = 0.5,  # a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.4, # curved edges is always a nice tough
     edge.arrow.size = 0.1) # arrow size smaller

```

# Without names

### Blue nodes are RU staff, yellow nodes are non-RU staff

```{r}
plot(net1,
     vertex.color = in_network$vcol, 
     vertex.label = NA,
     vertex.size = 5,  # we'll make them a bit smaller
     vertex.frame.color = "gray",  # we'll put a frame around it
     vertex.label.color = "black",  # not that ugly blue color for the names
     vertex.label.family = "Helvetica", # not a fan of times in figures
     vertex.label.cex = 0.5,  # a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.4, # curved edges is always a nice tough
     edge.arrow.size = 0.1) # arrow size smaller

```

# Simple descriptive statistics
```{r}
# number of nodes
vcount(net1)
# number of edges
ecount(net1)

# degree
degree(
  graph = net1,
  v = V(net1),
  mode = c("all", "out", "in", "total"),
  loops = TRUE,
  normalized = FALSE
)

# Closeness centrality measures how many steps is required to access every other vertex from a given vertex.
closeness(
  net1,
  vids = V(net1),
  mode = c("out", "in", "all", "total"),
  weights = NULL,
  normalized = FALSE,
  cutoff = -1
)


# Transitivity is the number of transitive triads divided by all possible transitive triads.
transitivity(net1, type = c("localundirected"), isolates = c("NaN", "zero"))

# Betweenness centrality is the amount of times a node is on the route of the shortest path from a node to another node
igraph::betweenness(net1, directed = FALSE)

# Standardization of betweenness centrality value
# ((value of betweenness node) - minimum value betweenness centrality) / maximum value betweenness centrality - minimum value betweenness centrality



```
