
# Setting up everything in the background
```{r}
rm(list = ls())


packages = c("RSiena", "tidyverse", "stringdist", "stringi", "dplyr", "stringr")

require("magrittr")
library(tidyverse)
library(stringr)
library(RSiena)
library(stringdist)
library(stringi)
library(dplyr)
library(stringr)

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}



# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019), wave2 = c(2020, 2021), wave3 = c(2022, 2023))) {

    publications <- list_publications %>%
        dplyr::bind_rows() %>%
        dplyr::distinct(title, .keep_all = TRUE)

    df_scholars %>%
        dplyr::filter(affil1 == affiliation | affil2 == affiliation) %>%
        dplyr::filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            dplyr::select(author) %>%
            lapply(str_split, pattern = "," )
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```


# Merging datasets

```{r, eval=FALSE}
total <- merge(df_soc, soc_df,by="gs_id")
view(total)

save(total,file="totaldata.Rda")

```




# getting the dataset ready
```{r}
df <- fload("./Data/Data_processed/totaldata.rda")
publications <- fload("./Data/Data_processed/20230621list_publications_jt.rda")


output <- f_pubnets(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2017, 2018, 2019), wave2 = c(2020, 2021, 2022)))

df_soc <- output[[1]]
df_network <- output[[2]]
```

## Matrices for different waves



### Wave 1
```{r}

library(igraph)
library(dplyr)

g1 <- graph.adjacency(df_network[[1]])
g1 <- get.edgelist(g1)

diag(g1) <- 0

edgelistwave1 <- data.frame(g1)

edgeswave1 <- as.matrix(edgelistwave1, rownames.force = TRUE)  # matrix because igraph wants that

# Remove rows with negative node IDs
edgeswave1 - min(edgeswave1) + 1


# Shift node IDs to start from 1
edgeswave1 <- edgeswave1 - min(edgeswave1) + 1

net1 <- graph_from_edgelist(edgeswave1, directed = TRUE)  # Notice the igraph function here!

plot(net1)

df_soc$RID <- 1:nrow(df_soc)

# saver way to get columns by name
rid_gender_column_names <- grep("RID|gender", names(df_soc), value = TRUE)
print(rid_gender_column_names)

onlyRU <- dplyr::left_join(edgelistwave1, df_soc[, rid_gender_column_names], by = c("X1" = "RID"))
onlyRU <- dplyr::left_join(onlyRU, df_soc[, rid_gender_column_names], by = c("X2" = "RID"))
```





### Wave 1
```{r, eval=FALSE}
library(igraph)
library(dplyr)


g1 <- graph.adjacency(df_network[[1]])
g1 <- get.edgelist(g1)

diag(g1) <- 0

edgelistwave1 <- data.frame(g1)

edgeswave1 <- as.matrix(edgelistwave1)  # matrix because igraph wants that
net1 <- graph_from_edgelist(edgeswave1, directed = TRUE)  # Notice the igraph function here!

plot(net1)



df_soc$RID <- 1:nrow(df_soc)


onlyRU <- left_join(edgelistwave1, df_soc[, c("RID", "gender.y")], by = c ("X1" = "RID"))
onlyRU <- left_join(onlyRU, df_soc[, c("RID", "gender.y")], by = c ("X2" = "RID"))
```

### wave 2
```{r}
g2 <- graph.adjacency(df_network[[2]])
g2 <- get.edgelist(g2)

# diag(g2) <- 0

edgelistwave2 <- data.frame(g2)

edgeswave2 <- as.matrix(edgelistwave2)
net2 <- graph_from_edgelist(edgeswave2, directed = TRUE)

plot(net2)

ecount(net2)

vcount(net2)




df_soc$RID <- 1:nrow(df_soc)

onlyRU2 <- left_join(edgelistwave2, df_soc[, c("RID", "gender.y")], by = c ("X1" = "RID"))
onlyRU2 <- left_join(onlyRU2, df_soc[, c("RID", "gender.y")], by = c ("X2" = "RID"))
```

## Making gender homophily variable

### Wave 1
```{r}
onlyRU$gendersimilar <- NA

onlyRU$gendersimilar[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "man"] <- 1 
onlyRU$gendersimilar[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "woman"] <- 1
onlyRU$gendersimilar[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "woman" ] <- 0
onlyRU$gendersimilar[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "man" ] <- 0

summary(onlyRU$gendersimilar)

onlyRU$percentageww <- NA

onlyRU$percentageww[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "man"] <- 0 
onlyRU$percentageww[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "woman"] <- 1
onlyRU$percentageww[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "man"] <- NA 
onlyRU$percentageww[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "woman"] <- NA

summary(onlyRU$percentageww)

onlyRU$percentagemm <- NA

onlyRU$percentagemm[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "man"] <- 1 
onlyRU$percentagemm[onlyRU$gender.y.x == "man" & onlyRU$gender.y.y == "woman"] <- 0
onlyRU$percentagemm[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "man"] <- NA 
onlyRU$percentagemm[onlyRU$gender.y.x == "woman" & onlyRU$gender.y.y == "woman"] <- NA

summary(onlyRU$percentagemm)


```

### Wave 2
```{r}
onlyRU2$gendersimilar2 <- NA

onlyRU2$gendersimilar2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "man"] <- 1 
onlyRU2$gendersimilar2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "woman"] <- 1
onlyRU2$gendersimilar2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "woman" ] <- 0
onlyRU2$gendersimilar2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "man" ] <- 0
onlyRU2$percentagemm2[onlyRU2$X1 == onlyRU2$X2] <- NA

summary(onlyRU2$gendersimilar2)

onlyRU2$percentageww2 <- NA

onlyRU2$percentageww2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "man"] <- NA 
onlyRU2$percentageww2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "woman"] <- 1
onlyRU2$percentageww2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "woman" ] <- NA
onlyRU2$percentageww2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "man" ] <- 0
onlyRU2$percentagemm2[onlyRU2$X1 == onlyRU2$X2] <- NA

summary(onlyRU2$percentageww2)


onlyRU2$percentagemm2 <- NA

onlyRU2$percentagemm2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "man"] <- 1 
onlyRU2$percentagemm2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "woman"] <- NA
onlyRU2$percentagemm2[onlyRU2$gender.y.x == "man" & onlyRU2$gender.y.y == "woman" ] <- 0
onlyRU2$percentagemm2[onlyRU2$gender.y.x == "woman" & onlyRU2$gender.y.y == "man" ] <- NA
onlyRU2$percentagemm2[onlyRU2$X1 == onlyRU2$X2] <- NA

summary(onlyRU2$percentagemm2)
```
## Getting descriptive statistics 

```{r}
summary(onlyRU$gendersimilar)

summary(onlyRU2$gendersimilar2)

ecount(net1)

vcount(net1)

ecount(net2)

vcount(net2)


```
```{r}
degreewave1 <- igraph::degree(net1)
degreewave2 <- igraph::degree(net2)
```

## Histogram of the number of degree in wave 1
```{r}
#par(mfrow=c(3,1))
hist(degreewave1, col="#99d6ff")
```
## Histogram of the number of degree in wave 2
```{r}
hist(degreewave2, col="#99d6ff")
```

## Network statistics wave 1 (dyad/triad/density)

```{r}
igraph::dyad.census(net1)

igraph::triad.census(net1)

igraph::edge_density(net1)
```

## Network statistics wave 2 (dyad/triad/density)

```{r}
igraph::dyad.census(net2)

igraph::triad.census(net2)

igraph::edge_density(net2)
```

## Transitivity wave 1
```{r}
igraph::transitivity(net1)
```

## Transitivity wave 2
```{r}
igraph::transitivity(net2)
```


## Descriptive statistics of the control variables

```{r}
hist(df$num_pubs, col="#99d6ff")

hist(df$career_age, col="#99d6ff")
summary(df$career_age)

hist(df$num_coauth, col="#99d6ff")
summary(df$num_coauth)


df$freq0 <- df$Freq                   
df$freq0[is.na(df$freq0)] <- 0


hist(df$freq0, col="#99d6ff")
summary(df$freq0)

```





## Network statistics - Still to do


outIso?

network-isolate effect, (isolateNet), the effect of ego having in-degree as well as out-degree zero, i.e., being a total isolate,snet i56(x, z) = I{x+i = xi+ = 0} ;
anti isolates effect, (antiIso), the effect of wishing to connect to others who otherwise would be a total isolate, i.e., have no incoming or outgoing ties, and wishing not to sever connections to others who thereby would become a total isolate,snet i57(x) = Pj I{x+j ≥ 1, xj+ = 0};


covariate-related similarity × reciprocity (simRecipX)

outdegree activity to same covariate (sameXOutAct) defined by the squared number of ties to those others who have the same covariate value as i


## RSiena Models


### Dependent variable 
```{r}
# Number of waves
length(df_network)
```

### Independent variables

