---
title: "First introduction to rSiena"
output:
  html_document:
    toc: true
    number_sections: true
author: "by: Stijn van Helmondt"
bibliography: references.bib

---

# Preparing R and loading packages and data

```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

packages = c("stringdist", "stringi", "tidyverse")

fpackage.check(packages)

df <- fload("./Data/Data_processed/20230621df_complete.rda")
publications <- fload("./Data/Data_processed/20230621list_publications_jt.rda")

# Number of unique publications in df
publications <- publications %>%
    bind_rows() %>%
    distinct(title, .keep_all = TRUE)

```

---

# Network based on publications

Understanding the data and decisions to make:

- What are the boundaries of the network?  
  + combination of university and discipline  
- Is it a directed, or undirected network?  
  + directed: first author is sending a tie to other authors  
- Is it a weighted or unweighted network?  
  + unweighted (no distinction between number ties within a specific time window, thus 0/1)  
- Time window of the data
  + One wave is three years
  + There are two waves
    - wave 1: 2018-2020
    - wave 2: 2021-2023

---

# Using the sociology department of the RU as example

## Selecting scholars
```{r}
df %>%
    filter(affil1 == "RU" | affil2 == "RU") %>%
    filter(discipline == "sociology") -> df_sel
fshowdf(df_sel)
```

## Preparing the network

```{r}
# Network based on publications
## We will use an adjacency matrix to store our network ties: the first author is sending a tie to other authors. We make the assumption that the composition of the network is stable!

networkw1 <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
networkw2 <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))

# Selecting publications

## wave1
publications %>%
    filter(gs_id %in% df_sel$gs_id) %>%
    filter(year %in% c(2018, 2019, 2020)) -> pub_w1

## wave2
publications %>%
    filter(gs_id %in% df_sel$gs_id) %>%
    filter(year %in% c(2021, 2022, 2023)) -> pub_w2

# Making a list for each publication

## wave1
pub_w1 <- str_split(pub_w1$author, ",") %>%
    # lowercase
lapply(tolower) %>%
    # Removing diacritics
lapply(stri_trans_general, id = "latin-ascii") %>%
    # only last name
lapply(word, start = -1, sep = " ") %>%
    # only last last name
lapply(word, start = -1, sep = "-")

# let us remove all publications with only one author
remove <- which(sapply(pub_w1, FUN = function(x) length(x) == 1) == TRUE)
pub_w1 <- pub_w1[-remove]

## wave2
pub_w2 <- str_split(pub_w2$author, ",") %>%
    # lowercase
lapply(tolower) %>%
    # Removing diacritics
lapply(stri_trans_general, id = "latin-ascii") %>%
    # only last name
lapply(word, start = -1, sep = " ") %>%
    # only last last name
lapply(word, start = -1, sep = "-")

# let us remove all publications with only one author
remove <- which(sapply(pub_w2, FUN = function(x) length(x) == 1) == TRUE)
pub_w2 <- pub_w2[-remove]

# Finding the position of the author --> with this knowledge we can fill the adjacency matrix

pubs <- pub_w1
for (ego in 1:nrow(df_sel)) {
    # which ego?
    lastname_ego <- df_sel$lastname[ego]
    # for all publications
    for (pub in 1:length(pubs)) {
        # only continue if ego is author of pub
        if (lastname_ego %in% pubs[[pub]]) {
            aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
            # only continue if ego is first author of pub
            if (aut_pot == 1) {
                # check all alters/co-authors
                for (alter in 1:nrow(df_sel)) {
                  # which alter
                  lastname_alter <- df_sel$lastname[alter]
                  if (lastname_alter %in% pubs[[pub]]) {
                    networkw1[ego, alter] <- networkw1[ego, alter] + 1
                  }
                }
            }
        }
    }
}

pubs <- pub_w2

for (ego in 1:nrow(df_sel)) {
    # which ego?
    lastname_ego <- df_sel$lastname[ego]
    # for all publications
    for (pub in 1:length(pubs)) {
        # only continue if ego is author of pub
        if (lastname_ego %in% pubs[[pub]]) {
            aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
            # only continue if ego is first author of pub
            if (aut_pot == 1) {
                # check all alters/co-authors
                for (alter in 1:nrow(df_sel)) {
                  # which alter
                  lastname_alter <- df_sel$lastname[alter]
                  if (lastname_alter %in% pubs[[pub]]) {
                    networkw1[ego, alter] <- networkw1[ego, alter] + 1
                  }
                }
            }
        }
    }
}
```

# Now all in one code

```{r}
f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}

```

# And a bit of testing

```{r}
outputUU_sociology <- f_pubnets(affiliation = "UU")
str(outputUU_sociology)

head(outputUU_sociology[[1]])
head(outputUU_sociology[[2]])

outputRU_sociology <- f_pubnets(affiliation = "RU")
str(outputRU_sociology)

head(outputRU_sociology[[1]])
head(outputRU_sociology[[2]])
```

