---
title: "First introduction to analyses rSiena"
output:
  html_document:
    toc: true
    number_sections: true
author: "by: Stijn van Helmondt"
bibliography: references.bib

---

# Preparing R and the packages etc.

```{r}
rm(list = ls())
```

## Goals of the workshop
- How to make an RSiena object ready to analyze.  
- Analyze a (very very) simple network evolution model

## Loading the packages

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
```

## Loading package RSiena and data

```{r}
packages = c("RSiena", "tidyverse", "stringdist", "stringi")

fpackage.check(packages)

df <- fload("./Data/Data_processed/20230621df_complete.rda")
publications <- fload("./Data/Data_processed/20230621list_publications_jt.rda")
```

# Analyzing the RU Sociology department

```{r}
output <- f_pubnets()
df_soc <- output[[1]]
df_network <- output[[2]]
```

## Defining dependent and independent variables

### Dependent variable
```{r}
# let us check the number of waves
length(df_network)

wave1 <- df_network[[1]]
wave2 <- df_network[[2]]
# let us put the diagonal to zero
diag(wave1) <- 0
diag(wave2) <- 0
# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
# put the nets in an array
net_soc_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))
# dependent
net <- sienaDependent(net_soc_array)
```

### Independent variable

```{r}
# gender
gender <- as.numeric(df_soc$gender == "female")
gender <- coCovar(gender)

mydata <- sienaDataCreate(net, gender)
```

#### Create effects structure

```{r}
myeff <- getEffects(mydata)
# effectsDocumentation(myeff)

```

#### Get initial description

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)
print01Report(mydata, modelname = "./results/soc_init")
```
## Specifying the model

```{r}
myeff <- includeEffects(myeff, isolateNet, inPop, outAct)  #we know that quite a lot of staff has not published with someone else

myeff <- includeEffects(myeff, sameX, egoX, altX, interaction1 = "gender")
```
## Estimating the model

```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate  further
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans, returnDeps = TRUE))
```

## Goodness of Fit

### Defining Goodness of Fit auilliary

```{r}
# see here: ?'sienaGOF-auxiliary'

# The geodesic distribution is not available from within RSiena, and therefore is copied from the
# help page of sienaGOF-auxiliary:

# GeodesicDistribution calculates the distribution of non-directed geodesic distances; see
# ?sna::geodist The default for \code{levls} reflects the usual phenomenon that geodesic distances
# larger than 5 do not differ appreciably with respect to interpretation.  Note that the levels of
# the result are named; these names are used in the \code{plot} method.
GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

### Applying SienaGOF

```{r}
gofi0 <- sienaGOF(ans, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0 <- sienaGOF(ans, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd <- sienaGOF(ans, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
# ?sienaGOF
```

### GOF plots initial model

```{r}
plot(gofi0)
```

### Relative Importance 

---

Until now, the interpretation of estimated effects in our SIENA models has been limited to testing their statistical significance, which determines whether an effect plays a role in the evolution of the network (using t-statistics). But we do not yet know how these effects fare against each other.

There are four issues when extrapolating the size of estimated parameters to their relative importance in SIENA models:

- Explanatory statistics have different scales (e.g., one micro-step may increase the number of reciprocated ties by at most 1 but may result in up to 2(N-2) new transitive triplets).

- Explanatory variables are often correlated, making it difficult to establish causality (e.g., a tie abridging a two-path may yield a new transitive triplet, while at the same time, a reciprocated tie).

- Multiple and complex choice sets exist, where network effects influence the probabilities of several alternative choices, and these effects are themselves influenced by a combination of several effects. This interdependence makes it challenging to assess the individual contribution of each effect to actor decisions.

- The data undergoes substantial unobserved changes over time, and the size of parameter estimates is strongly dependent on the structure of the evolving network data. The absence of certain network configurations can render specific effects irrelevant in decision-making processes at certain points in time (e.g., if an ego has no incoming ties, he has no opportunity to reciprocate a tie, making that the reciprocity effect cannot influence his decision).

This is where the concept of ‘Relative Importance’ (RI) measures comes into play. This measure reflects the extent that estimated model parameters affect change probabilities in network decision probabilities. They should be interpreted as the influence of effects on network changes relative to one another. The importance of an effect is estimated based on the extent to which network micro-steps would have differed if this effect were to be omitted. Probabilities for tie changes from the perspective of each actor are calculated using the fitted model parameters. Subsequently, each parameter is fixed to 0 and the change probabilities are recalculated. The influence of an effect on network (or: behavior) micro-steps is evaluated based on the magnitude of the difference in the distribution of change probabilities with the particular effect present versus absent. These differences are normalized so that their sum is 1 for each actor, and subsequently averaged across actors.


### Siena RI

```{r}
# get parameters
theta.eval <- ans$theta
# and effects
myeff.eval <- ans$effects

# use sienaRI()
RI <- sienaRI(data = mydata, theta = theta.eval, algorithm = myAlgorithm, effects = myeff.eval)

# Plotting
plot(RI, addPieChart = TRUE)
```

